---
title: "Take Home Final"
subtitle: "Due 07/18 @ 11:59pm"
author: "WRITE YOUR NAME HERE"
output: html_document
---

## README

Complete the following questions below and show all work.
You must submit a compiled html file with all code and answers in this document on Canvas.
Make sure to write your name above where it says "WRITE YOUR NAME HERE".
Include it within the quotation marks.

The total amount of points possible is 110.
You can find a table at the bottom of the document 
that denotes how points are allocated to each question.
Following previous guidelines, stay within the `tidyverse`, `broom`, and `here` packages.
You may not use other packages, although base R functions are allowed.
All code must be placed in code blocks and all regression estimates must be
presented in a `broom` tidy format.

To use broom after a regression, simply pipe the lm( ) into broom::tidy( ) as in the example below:

```{r, eval = FALSE}
lm(INCTOT ~ AGE, data = ipums_tbl) %>%
  broom::tidy()
```

__Style Points__.
Good code should be easy to read and presentation is an important skill.
10 points on the exam are dedicated for style points.
Style points are determined by the following criteria:

- Your submission is organized and follows the directions

- Your written answers are clear

- Code is concise and easy to read.
It follows the principles of the tidyverse as introduced in koan assignments

- Plots are labelled and clearly presented

_Note_: 
Before answering the following questions, make sure to compile the document to make sure that everything is working properly.
You can compile the document by clicking on the **knit** option on the bar above. 
As you answer each question, compile after each question to make sure you are not making any errors.
If you are experiencing an error, read the error carefully.
One tactic is to selectively delete questions/sections until the report compiles.
This will help you narrow down where the coding error is occurring.

#### Data

The data set used on this exam come from the 2022 American Community Survey (ACS), which was downloaded from [IPUMS](https://ipums.org/).
At the bottom of this document, you will find a table codebook that describes the variables in the data set.

#### Integrity

If you are suspected of cheating, you will receive a zero—for the assignment and possibly for the course.
Collaboration is not allowed.
All work must be your own.
If you have any questions about what constitutes cheating, please ask.

## Questions

### Setup

__Q01__. Use the following code block to load the data.
It should just work, so don't mess with it.
Compile the document and proceed to the next question.

```{r, include=FALSE}
# Install the 'here' package if it is not already installed
if (!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}

# Load packages
library(here)
library(tidyverse)
library(broom)

# Define the path to the data file using the 'here' package
ipums_tbl = here("data", "ipums_2022.csv") %>%
  readr::read_csv() 
```

---

# Step 0: Define a research question

In this assignment, you will be looking at the return to education for 
various different types of degrees and college majors. 

Our main research question will be:
___Do returns to educational attainment differ by college major?___

# Step 1: Observe the Data 

### __Q01: Take a look at the ipums dataset. Describe it's structure in words.__ 

  - How many observations are there? 

  - Is it cross-sectional, time-series, or panel data?
  Who/what does each observation (row) describe?

  - What qualitative information is contained in this dataset? 

  - What quantitative information is contained in this dataset? 

*Use the codebook at the end of this document to help you answer this question.*

### __Q02: How many people who received a degree in economics are now unemployed?__

```{r}
# Write your code here
```

### __Q03: How does the unemployment rate for economists compare to other types of majors?__

```{r}
# Write your code here
```

### __Q04: How does the unemployment rate for people who's highest degree is a bachelors degree compare to those for whom their highest degree is only a high school diploma?__

```{r}
# Write your code here
```

### __Q05: Now take a closer look at the INCTOT variable. Describe the shape of the distribution of personal incomes within this dataset.__

  - Label your axis and include a title and as explain in words what your 
  figure teaches us about personal incomes.

```{r}
# Write your code here
```

### __Q06: Create a new variable for log incomes instead. How does the shape of the distribution change?__

```{r}
# Write your code here
```

### __Q07: Now plot the INCTOT together with years of school attainment__

**WARNING: A scatterplot with all 100,000 observations will take a very long time to generate and will be hard to see the trend! You should use an alternative like a bin scatter plot, a geom_hex, or subset to a smaller sample**

```{r}
# Write your code here
```

  - Can you see any visual evidence of a relationship?

### __Q08: Based on your graphs, are there any considerations of which transformations of variables to use in your regressions? __
  
# Step 3: Analyze the Data

### __Q09: Regress your transformation of the income variable on school attainment. Report and interpret your estimates in the context of the returns to education __
  
```{r}
# Write your code here
```

### __Q10: Does there seem to be any evidence of heteroskedasticity in this relationship? Plot the residuals from your previous regression at different levels of school attendance and interpret. __

```{r}
# Write your code here
```

### __Q11: What are some potential omitted variables in the relationship between school attainment and income which could bias your regression estimates? __

  - How might these variable be related to *both* income and school attainment?
  
  - Choose one variable and argue which direction (positive or negative) of bias 
  likely would result from omitting that variable.

### __Q12: Now regress income on school attainment, age, and on whether the person received a degree in economics__

```{r}
# Write your code here
```

  - How does the estimate on school attainment change? 
  
  - How does your interpretation change?

### __Q13: Regress income on school attainment and the interaction between school attainment and the degree category of your choice. __

```{r}
# Write your code here
```

# Step 4: Discuss your conclusions

### __Q14: In this dataset, does it seem like some degrees have better returns to lifetime income than others?__

  - Make sure to cite your estimate results and explain why they support 
  your claim.
  
### __Q15: What are some reasons why your estimates may not be enough to answer the question of whether degree type actually has a *causal* effect on the returns to education? __

<br>

---

## Points

| Question             | Points        |
|----------------------|---------------|
| Q01                  | 8             |
| Q02                  | 2             |
| Q03                  | 4             |
| Q04                  | 4             |
| Q05                  | 6             |
| Q06                  | 6             |
| Q07                  | 6             |
| Q08                  | 2             |
| Q09                  | 8             |
| Q10                  | 6             |
| Q11                  | 8             |
| Q12                  | 10            |
| Q13                  | 10            |
| Q14                  | 10            |
| Q15                  | 10            |
| Style Points         | 10            |
| Total                | 110           |



---

## Codebook

| Variable     | Description                                                                           |
|--------------|---------------------------------------------------------------------------------------|
| STATEFIPS    | State name                                                                            |
| AGE          | The individual's age (in years)                                                       |
| INCTOT       | Personal pre-tax income (in 2022 dollars)                                             |
| school       | Estimated number of years of education attainment (based on courser measure in ipums) |
| i_unemployed | Binary indicator for whether the individual is unemployed                             |
| i_business   | Binary indicator for whether the individual received a bachelors degree in business   |
| i_education  | Binary indicator for whether the individual received a bachelors degree in education  |
| i_engineering| Binary indicator for whether the individual received a bachelors degree in engineering|
| i_medicine   | Binary indicator for whether the individual received a bachelors degree within 'Medical and Health Sciences and Services' |
| i_socialsci  | Binary indicator for whether the individual received a bachelors degree in a social science |
| i_economics  | Binary indicator for whether the individual received a bachelors degree in economics  |
| i_hsdiploma  | Binary indicator for whether the individual received at least a high school diploma   |
| i_ged        | Binary indicator for whether the individual received at least a GED                   |
| i_associates | Binary indicator for whether the individual received at least an associates degree    |
| i_bachelors  | Binary indicator for whether the individual received at least a bachelors degree      |
| i_masters    | Binary indicator for whether the individual received at least a masters degree        |
| i_doctorate  | Binary indicator for whether the individual received a doctoral degree                |
