\documentclass[12pt]{exam}
\usepackage{amsthm}
\usepackage{libertine}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{multicol}
\usepackage[shortlabels]{enumitem}
\usepackage{siunitx}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{tikz}
\usepackage[T1]{fontenc}  % Ensure proper encoding
\usepackage{titlesec}
\usepackage{booktabs}

\CorrectChoiceEmphasis{\color{red}\bfseries\itshape}

\newcommand{\class}{\fontfamily{lmss}\selectfont \textbf{EC 320}} % This is the name of the course 
\newcommand{\examnum}{\fontfamily{lmss}\selectfont \textbf{Problem Set 03}} % This is the name of the assignment
\newcommand{\examdate}{\textbf{Due 07/11 @ 11:59pm}} % This is the due date
\newcommand{\timelimit}{}
\addpoints % Allows to add points up to include table at end of document

\pagestyle{headandfoot} % Fancy equivalent for exam documentclass
\firstpageheadrule % Horizontal bar in first page
\runningheadrule % Horizontal bar in rest of pages
\firstpageheader{\class}{\examnum}{\fontfamily{lmss}\selectfont  \examdate} % 1st page header content
\firstpagefooter{}{Page \thepage}{} % 1st page footer content
\runningheader{\class}{\examnum}{\fontfamily{lmss}\selectfont  \examdate} % Rest of pages header content
%\runningfooter{Points earned: \makebox[1in]{\hrulefill} / \pointsonpage{\thepage} points}{}{\thepage\ of \numpages} % Rest of pages footer content
\bracketedpoints % Puts points per question inside brackets instead of parenthesis 

\titleformat{\section}
  {\normalfont\large\bfseries\sffamily}{\thesection}{1em}{}

\titleformat{\subsection}
  {\normalfont\normalsize\bfseries\sffamily}{\thesection}{1em}{}

\begin{document}
\fontfamily{lmss}\selectfont

Suppose the true data generating process in the examples below are of the form 

\begin{align*}
    y_{i} = \beta_{0} + \beta_{1} x_{i} + \beta_{2} o_{i} + u_{i},
\end{align*}

where $x_{i}$ and $o_{i}$ are exogenous such that $\mathbb{E}[u_{i}|o,x] = 0$.

However, $o$ is a variable that cannot be observed and so instead, we run the model: 

$y_{i} = \beta_{0} + \beta_{1} x_{i} + \varepsilon_{i}$ where $\varepsilon_{i} = \beta_{2} o_{i} + u_{i}$

Recall that we can sign the bias by:

\begin{align*}
    \mathbb{E}[\hat{\beta}_{1}] = \beta_{1} + \beta_{2} \dfrac{Cov(X_{i},Z_{i})}{Var(X_{i})}
\end{align*}

\begin{questions}

\subsection*{Causal Effect of Migration on Earnings}

\question %01
Suppose we took survey data that included people's earnings and whether they have recently moved to a new city or not, and we estimated this model:

\begin{align*}
    Earnings_{i} = \beta_{0} + \beta_{1} Migration_{i} + \varepsilon_{i}
\end{align*}

Consider the unobservable variable absorbed in $\varepsilon_{i}$ which is $o_{i} = ambition_{i}$.

\begin{parts}
    \part How does $Ambition$ correlate with $Earnings$? That is, would you expect $\beta_{2}$ to be positive, negative or 0? 
    \vspace*{\stretch{1}}
    \part How does $Ambition$ correlate with $Migration$? 
    That is, would you expect 
    
    $Cov(migration,ambition)$ to be positive, negative or 0?
    \vspace*{\stretch{1}}
    \part We have to omit $ambition$ from the model because it is not observable. Considering your answers to parts (a) and (b), we should expect that $\hat{\beta}_{1}$ will be biased (circle one): up / down / not at all. 
    \vspace*{\stretch{1}}
\end{parts}

\newpage 

\subsection*{Causal Effect of Friends of the Opposite Sex on a High School Student's GPA}

\question %02
Suppose we took survey data that included high school student's GPAs and the number of friends of the opposite sex they have, and we estimated the following model:

\begin{align*}
    GPA_{i} = \beta_{0} + \beta_{1} \text{Opposite Sex Friends}_{i} + \varepsilon_{i}
\end{align*}

Consider the unobservable variable absorbed in $\varepsilon_{i}$ which is $o_{i} = \text{strict parents}_{i}$. 

\begin{parts}
    \part How does \emph{Strict Parents} correlate with $GPA$?
    That is, would you expect $\beta_{2}$ to be positive, negative or 0?
    \vspace*{\stretch{1}}
    \part How does \emph{Strict Parents} correlate with \emph{Opposite Sex Friends}? 
    That is, would you expect $Cov(\emph{Opposite Sex Friends}, \emph{Strict Parents})$ to be positive, negative, or 0?
    \vspace*{\stretch{1}}
    \part We have to omit \emph{Strict Parents} from the model because it is not observable. 
    Considering your answers to parts (a) and (b), we should expect that $\hat{\beta}_{1}$ will be biased (circle one): up / down / not at all.
    \vspace*{\stretch{1}} 
\end{parts}

\newpage 

\question %03

For the questions below, suppose you estimate the following model:

\begin{align*}
    weight = 0 - 140 \cdot female + 30 \cdot height + 20 \cdot height \times female + u 
\end{align*}

\begin{parts}
    \part How much would you predict a 6 foot male weighs?
    \vspace*{\stretch{1}}
    \part How much would you predict a 6 foot female weighs?
    \vspace*{\stretch{1}}
    \part What is the estimated impact of a one-unit (one foot) increase in height for a male individual?
    \vspace*{\stretch{1}}
    \part What is the estimated impact of a one-unit (one foot) increase in height for a female person?
    \vspace*{\stretch{1}}
    \part Which group has the steeper slope when it comes to how height $(x)$ impacts weight $(y)$?
    \vspace*{\stretch{1}}
\end{parts}

\end{questions}

\end{document}