---
name: install R
---

## Step 1: Grab a Computer

Good news! 
If you are reading this, you are most likely on a computer already. 
Great work so far.

The good news about R is that it works on any operating system. 
I personally use it on both Mac and Windows. 

---

## Step 2: Download and Install R

Go to https://cran.r-project.org/ and follow the instructions to download R for your device. 

---

## Step 3: Install RStudio

Go to https://posit.co/download/rstudio-desktop/ and click the blue button that says step 2: install Rstudio Desktop.
Follow the instructions to complete the installation. 

---

## Step 4: Open RStudio and Install Some Packages

Run these lines of code in your console to make sure you have the `tidyverse` installed and attached to your current session. 

```{r}
#| eval: false
#| echo: true

install.packages("tidyverse", dependencies = TRUE)
library(tidyverse)
```

<br>

An important shortcut to installing and using packages is a package called `pacman`. 
It has a function `p_load()` which functions as both `install.packages()` and `library()` above.

```{r}
#| eval: false
#| echo: true

install.packages("pacman")
```

Use `p_load` to install the next set of packages

---

## Step 4a: Install More Packages

A former PhD Student (Colleen O'Briant) in the Econ Department created a lovely package that serves as an alternative set of beginner friendly help docs. 

Run this code to install it:

```{r}
#| eval: false
#| echo: true

library(pacman)
p_load(Rcpp, devtools) # Note you can do several packages at once

install_github("cobriant/qelp")
```

To test everything worked, now run:

```{r}
#| eval: false
#| echo: true 

?qelp::install.packages
```

If everything went right, the help docs she wrote on the function `install.packages` should pop up. 

::: {.small}
> [Note:]{.note} It is not a complete library of every function, but many of the basic ones we will use are included. 
:::