---
title: "dplyr-sol"
author: "Jose Rojas"
date: "2025-07-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Solutions

```{r, include = FALSE}
library(tidyverse)
library(gapminder)
view(gapminder)
```

## filter()

### Q01: Filter gapminder for all the observations from Europe in 2007

```{r}
gapminder %>% 
  filter(continent == "Europe" & year == 2007)
```

### Q02: Filter gapminder for all the observations where lifeExp was exactly equal to 68 years old.

```{r}
gapminder %>% 
  filter(lifeExp == 68)
```

## Logical Operators

### Q03: Filter gapminder for a short list of the richest countries in Asia.

```{r}
gapminder %>%
  filter(continent == "Asia", gdpPercap > 30000)
```

### Q04: Filter gapminder for observations *IN* the United States, Germany, and Brazil.\

```{r}
gapminder %>%
  filter(country %in% c("United States","Germany","Brazil"))
```

## select()

### Q05: Select the last 3 variables of 'gapminder' by name

```{r}
gapminder %>%
  select(lifeExp, pop, gdpPercap)
```

## mutate()

### Q06: Use 'mutate()' to create a new variable 'pop_in_thousands'. So if the observation has 'pop' = 97,462, then 'pop_in_thousands' will be 97.462.

```{r}
gapminder %>%
  mutate(pop_in_thousands = pop/1000)
```

## summarize()

### Q07. Take 'gapminder', filter for only observations in Africa, and summarize to find the:

  - median life expectancy,
  - median population, and
  - median GDP per capita.
  
```{r}
gapminder %>%
  filter(continent == "Africa") %>%
  summarise(
    median_life = median(lifeExp),
    median_pop = median(pop),
    median_gdp = median(gdpPercap)
  )
```
  
  
### Q08. Take 'gapminder', add a new column (mutate) for the total gdp, and summarize to find the mean and median total gdp.

```{r}
gapminder %>%
  mutate(total_gdp = sum(gdpPercap)) %>%
  summarise(
    median_gdp = median(total_gdp),
    mean_gdp = mean(total_gdp)
  )
```


## group_by()

### Q08. Take 'gapminder', filter for only observations in Africa, and summarize to find the median life expectancy, population, and GDP per capita *for each country*.

```{r}
gapminder %>%
  filter(continent == "Africa") %>%
  group_by(country) %>%
  summarise(
    median_life = median(lifeExp),
    median_pop = median(pop),
    median_gdp = median(gdpPercap)
  )
```


### Q09. Summarize 'gapminder' to find the mean GDP per capita for each continent, for each year (use 2 variables inside 'group_by').


```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarise(
    mean_gdp = mean(gdpPercap)
  )
```

## count()

### Q10. How many observations are there from each country?

```{r}
gapminder %>%
  count(country)
```

## arrange() and slice()

### Q11. Take all the observations in Asia and sort them from the lowest life expectancies to the highest.

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  arrange(lifeExp)
```

### Q12. Take all the observations in Asia and sort them from the highest life expectancies to the lowest.

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  arrange(desc(lifeExp))
```

### Q13. Select row 853 to row 864.

```{r}
gapminder %>%
  slice(853:864)
```

### Q14. Take all the observations in Asia and return the five with the highest life expectancies (hint: use arrange()).

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  arrange(desc(lifeExp)) %>%
  slice_head(n = 5)
```

### Q15. Which (one) country has the highest life expectancy in each continent? (hint: use arrange() and also group_by()).

```{r}
gapminder %>%
  group_by(continent) %>%
  arrange(desc(lifeExp)) %>%
  slice_head(n = 1)
```
