---
title: "tibbles-lm-sol"
author: "Jose Rojas"
date: "2025-07-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
```

# Solutions

## Building a Tibble

Now take some data for Canada:
In 1957, GDP per capita in Canada was $12,490 and life expectancy was 70.0 years.
In 1977, GDP per capita in Canada was $22,091 and life expectancy was 74.2 years.
In 1997, GDP per capita in Canada was $28,955 and life expectancy was 78.0 years.

### Q01. Create a tibble that holds the canada_data

```{r}
canada_data <- tibble(
  year      = c(1957,1977,1997),
  gdpPercap = c(12490,22091,28955),
  lifeExp   = c(70.0,74.2,78.0)
)

canada_data
```

### Q02. How many columns does us_data have?

```{r}
us_data <- tibble(
  year = c(1957, 1977, 1997),
  gdpPercap = c(14847, 24073, 35767),
  lifeExp = c(69.5, 73.4, 76.8)
)
```

```{r}
ncol(us_data) == 3
```

### Suppose you want to append a row (or several) to your dataset. 
You can do that using the function 'add_row()'.
It takes the tibble as the first argument and then variable names and values to construct the new row. 
Suppose we got a new observation (row) about Canada: 
In 2007, GDP per capita in Canada was $36,319 and life expectancy was 80.7 years.

```{r}
add_row(
  canada_data,
  year = 2007,
  gdpPercap = 36319,
  lifeExp = 80.7
)
```

## Estimates, Standard Errors, and R-Squared

### Q03. Simple Regression: How much can a $1 increase to a country's GDP per capita be expected to boost that country's life expectancy? Recall the lm() syntax is: lm(y ~ x, data = gapminder)

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap, data = .)
```

A $1 increase to a country's GDP per capita can be associated with a 0.0007649 year increase in life expectancy.


### Q04. Pipe the simple regression from question 1 into `broom::tidy(conf.int = TRUE)`. 
Is the coefficient on `gdpPercap` greater than 0 at the 1% level?

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap, data = .) %>%
  broom::tidy(conf.int = TRUE)
```

The coefficient on `gdpPercap` is greater than 0 at the 1% level.

### Q05. To see the regression's r-squared, pipe the lm() call into `broom::glance()`.

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap, data = .) %>%
  broom::glance()
```

This model explains __% of the variation in lifeExp.

## Fitted Values and Residuals

### Q06. Pipe the simple regression into fitted.values() to get a vector of length 1704 (the same length as the number of rows of gapminder).

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap, data = .) %>%
  fitted.values()

```

### Q07. Pipe the simple regression into residuals() to get another vector of length 1704. 
Recall that the fitted values plus the residuals will be equal to the observed values for lifeExp.

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap, data = .) %>%
  residuals()
```

## Transforming Variables

### Q08. Take the simple regression we've been working on and apply a log transformation to both the dependent variable and the explanatory variable by wrapping the variable names in the function log():
lm(log(y) ~ log(x), data = gapminder).

```{r}
gapminder %>%
  lm(log(lifeExp) ~ log(gdpPercap), data = .) %>%
  broom::tidy(conf.int = TRUE)
```

### Q09. Instead of applying log() to both the dependent and explanatory variable, apply it only to the explanatory variable. Does the R-squared improve?

```{r}
gapminder %>%
  lm(lifeExp ~ log(gdpPercap), data = .) %>%
  broom::glance()
```

### Q10. Instead of doing a log transformation, try taking the square of gdpPercap. 
You'll need to wrap the transformation in I(), which "inhibits the interpretation" of arithmetic operators like "+", "-", "*", and "^" in formulas. So the formula will look like: lm(y ~ I(x^2), data = gapminder).
What's the R-squared now? 
Does a log transformation, square transformation, or no transformation at all of gdpPercap seem to be the best fit?

```{r}
gapminder %>%
  lm(lifeExp ~ I(gdpPercap^2), data = .) %>%
  broom::glance()
```


### Q11. If you wanted to omit the intercept term, you can do that like this: 
lm(y ~ 0 + x, data = gapminder). 
Take the simple regression and omit the intercept.

```{r}
gapminder %>%
  lm(lifeExp ~ 0 + gdpPercap, data = .) %>%
  broom::tidy(conf.int = TRUE)
```

### Q12. Take the simple regression we've been working on and add a second explanatory variable: 'year'.

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap + year, data = .) %>%
  broom::tidy(conf.int = TRUE)
```


Every year, a country's life expectancy is expected to increase by 0.2389 years.

### Q13. Take out 'year' from the model and add 'continent'. 
This variable is different from the numerical variables we've been working with (lifeExp, gdpPercap, and year are all numeric). 
'continent' is a "factor" variable, which means the data is categorical rather than numeric. 
When you add 'continent', lm() lets different continents have different intercepts.

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap + continent, data = .) %>%
  broom::tidy(conf.int = TRUE)
```

What happened to Africa? One level has to be omitted to avoid the dummy variable trap. 
So the intercept is the OLS prediction for the intercept for Africa. 
It's saying that if a country in Africa has a GDP per capita of $0 (which is nonsense), the life expectancy of the people in that country is expected to be 47.9 years. 
The intercept for a country in the Americas is 47.9 + 13.6 = 61.5 years.

Our model predicts that a country in Asia with $0 GDP per capita will have a life expectancy of 47.88 + 8.65 years.

### Q14. Estimate a model of lifeExp where the explanatory variable is the interaction between 'gdpPercap' and 'continent'. 
Hint: an interaction between x and z is done like this: lm(y ~ x:z, data = gapminder)

```{r}
gapminder %>%
  lm(lifeExp ~ gdpPercap:continent, data = .) %>%
  broom::tidy(conf.int = TRUE)
```

According to our model, a $1 larger GDP per capita can be associated with a 0.00101 year higher life expectancy in Europe.








