---
title: "vectors-pipes-sol"
author: "Jose Rojas"
date: "2025-07-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
```

# Solutions

## Create a vector with `c()`

### Q01. Make the first element of this numeric vector '6'. 

```{r}
c(6, 4, 5, 2, 3)
```

### Q02. Make the third element of this character vector 'economics'. 
Note that quotes shouldn't be used with numbers, but should be used with character strings.

```{r}
c("apple", "banana", "economics")
```

## Calculations on vectors

### Q03. Add these two numeric vectors.

```{r}
c(6, 3, 2) + c(3, 2, 1) == c(9, 5, 3)
```

### Q04. Find the minimum of this vector.


```{r}
min(c(6, 3, 2)) == 2
```

### Q05. Find the sum of all elements of this vector.

```{r}
sum(c(6, 3, 2)) == 11
```

### Q06. Multiply a scalar and a vector.\

```{r}
100 * c(6, 3, 2) == c(600, 300, 200)
```

### Q07. Divide two vectors.

```{r}
c(6, 3, 2) / c(2, 3, 1) == c(12, 9, 2)
```

## Length

### Q08.  Create a vector of length 5. It can be a character vector or a numeric vector.

```{r}
length(c(1:5)) == 5
```

### Q09. Create a random character vector that draws "heads" or "tails".

```{r}
sample(c("heads","tails"), size = 5, replace = TRUE)
```

## Pipes

```{r}
us_data <- tibble(
  year      = c(1957, 1977, 1997, 2017),
  gdpPercap = c(14847, 24073, 35767, 60062),
  lifeExp   = c(69.5, 73.4, 76.8, 78.5)
)
```

### Q10. Your turn: apply 'print' to us_data using a pipe.

```{r}
us_data %>% print()
```



















